---
title: "cfp model prototype"
author: "jean-mathieu vermosen"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
Sys.setenv(RETICULATE_PYTHON = '/opt/miniconda/bin/python')
knitr::opts_chunk$set(echo = TRUE)

library(reticulate)
```

## the Chen Feng Palomar model 

This document present the Chen-Feng-Palomar model in python

```{python}
import sys
sys.path.append('/home/jvermosen/git/otos/lib/Release/python')

import otos.cfp
import numpy as np
import datatable as dt
from datatable import f
```

the cfp model is specified by the following system:

$$
\begin{aligned}
y_t &= x_t \cdot c + \epsilon_t \\
x_{t+1} &= A_tx_t + \gamma_t
\end{aligned}
$$
with $\epsilon_t\sim \mathcal N(0, r)$ and $\gamma_t\sim\mathcal N(0, Q_t)$. Furthermore, the matrix $A_t$ is such that:

$$
\begin{aligned}
A_t=\left[
\begin{array}{cc}
\eta_t& 0\\
0 & \nu_t
\end{array}
\right]
\end{aligned}
$$

```{python}
p = otos.cfp.parameters()
p.a_eta = 0.3
p.a_mu  = 1.0
p.s_eta = 1e-15
p.s_mu  = 1e-15
p.r     = 1e-15
p.pi    = [0.3, 1.0]
p.sigma = [1e-15, 1e-15]
p.psi   = [1.0, 1.0, 1.0, 1.0]
```

* simulate according to the passed parameters

```{python}
import numpy as np
np.random.seed(1234)
data = otos.cfp.simulate(p, size=1000)
```

# create model object and compute the kalman filter/smoother of the time series

```{python}
model = otos.cfp.model(p)
flt = model.filter(data)
smt = model.smoother(data)
```

# apply the em algorithm to retrieve the parameters

```{python}

# dummy
model = otos.cfp.model(p)
fit = model.emax(data, 10000, 1e-15)
```

# plot 

```{python}

import plotly, json
import plotly.graph_objects as go
from plotly.subplots import make_subplots

pl = make_subplots(rows = 1, cols = 1, shared_xaxes=True)

pl = pl.add_trace(go.Scattergl(y = data, name = 'simulation'
    ), row=1, col=1)
pl = pl.add_trace(go.Scattergl(y = flt, name = 'filter'
    ), row=1, col=1)
pl = pl.add_trace(go.Scattergl(y = smt, name = 'smoother'
    ), row=1, col=1)
    
plotly.io.write_html(pl, file='cfp_fig1.html', auto_open=False)
```

```{r echo=FALSE}
htmltools::includeHTML("cfp_fig1.html")
```

```{python}

np.random.seed(1234)

p = otos.cfp.parameters()
p.a_eta = 0.93970117964
p.a_mu  = 0.617656427391
p.s_eta = 0.00611157182105
p.s_mu  = 0.100157446738
p.pi    = [-0.0377453676475, 0.637412180444]
p.sigma = [9.49012643982e-05, 0.00182247858807]
p.r     = 0.0670554645138
p.psi   = [ 
           15.3322974898, 14.8914481022, 14.8960953411
         , 14.7328251327, 14.717101214 , 14.600429107 , 14.5878141379
         , 14.51710026  , 14.4863107795, 14.381950856 , 14.3423713871
         , 14.2429064605, 14.2109154215, 14.1635064751, 14.2339929961
         , 14.1522787054, 14.1947867672, 14.1869497863, 14.4385168913
         , 14.3792550185, 14.4077635371, 14.4649051166, 14.624926768
         , 14.7170306255, 15.0749366985, 15.9959231225 
        ]

md = otos.cfp.model(p)
df = dt.Frame(sim=otos.cfp.simulate(p, size=1000))
df = df[:, f[:].extend({'foo': md.predict(df[:, 'sim'], 10)})]
```